```{r setup, include=FALSE,error=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=FALSE,message=FALSE)
```

## Introduction

## Prerequsites 

```{r}
library(tidyverse)
library(sf)
library(terra)
library(kableExtra)
```

Load in the CPAD_2023a_SuperUnits.shp and the gHM_masked.tif files.

```{r}
cpad_super <- st_read(here::here("data", "discussion_data", "CPAD_2023a_SuperUnits.shp"), quiet=TRUE) %>%
  sf::st_transform("EPSG:4326") %>%
  janitor::clean_names() %>%
  mutate(ID = row_number())

ghm <- rast(here::here("data", "discussion_data", "gHM_masked.tif")) %>%
  project("EPSG:4326")
```

```{r}
plot(cpad_super["suid_nma"])
plot(ghm)
```

## Exercises

1. Let's rasterize the cpad dataset a few times! This is necessary to do zonal statistics!

```{r}
id_rast = rasterize(cpad_super, ghm, field = "suid_nma")

plot(id_rast)

access_rast = rasterize(cpad_super, ghm, field = "access_typ")

plot(access_rast)
```

2. Let's say you're interested in finding protected areas with intermediate levels of human modification. How could you do this?

a) masking and zonal statistics

```{r}
rcl = matrix(c(-Inf, 0.4, NA, 
               0.6, Inf, NA), ncol = 3, 
             byrow = TRUE)

mod_dis = classify(ghm, rcl = rcl)

plot(mod_dis)
```

Say you want to make this more complicated, and you want two groups?

```{r}
rcl2 = matrix(c(-Inf, 0.4, NA, 
                0.4, 0.5, 1,
                0.5, 0.6, 2, 
               0.6, Inf, NA), ncol = 3, 
             byrow = TRUE)

mod_dis2 = classify(ghm, rcl = rcl2)

plot(mod_dis2)
```

Or you just want to group them generally! 

```{r}
rcl3 = matrix(c(-Inf, 0.25, 1, 
                0.25, 0.5, 2,
                0.5, 0.75, 3, 
               0.75, Inf, 4), ncol = 3, 
             byrow = TRUE)

mod_dis3 = classify(ghm, rcl = rcl3)

plot(mod_dis3)
```

Now let's use zonal statistics to get a mean

```{r}
ghm_zonal = zonal(mod_dis, id_rast, fun = "mean", na.rm = TRUE)

head(ghm_zonal)
```

b) combining and filtering 

```{r}
cpad_ghm_values = terra::extract(x = ghm, y = cpad_super)
```

Summarize the mean of human modification within each protected area

```{r}
cpad_ghm_summary = cpad_ghm_values %>% 
  group_by(ID) %>% 
  summarise(gHM_mean = mean(gHM))
```

Join this summary with the protected area database. Plot these values

```{r}
cpad_ghm = full_join(cpad_super, cpad_ghm_summary)
```

Now we can filter! 

```{r}
cpad_ghm_sub = cpad_ghm %>% 
  st_drop_geometry() %>% 
  filter(gHM_mean < 0.6 & gHM_mean >= 0.4) %>% 
  distinct() %>% 
  select(suid_nma, gHM_mean) %>% 
  mutate(method = "Method 2")
```

3. These do give slightly different results however... can you identify the difference? 

```{r}
ghm_zonal = ghm_zonal %>% 
  filter(!is.na(gHM)) %>% 
  mutate(method = "Method 1")

method_sum = full_join(ghm_zonal, cpad_ghm_sub)

method_count = method_sum %>% 
  group_by(suid_nma) %>% 
  summarize(count = as.factor(n())) %>% 
  full_join(method_sum)
```

```{r}
ggplot(method_count) +
  geom_jitter(aes(method, suid_nma, color = count, alpha = count), height = 0) +
  theme_minimal() +
  scale_color_manual(values = c("red", "darkgrey")) +
  scale_alpha_manual(values = c(1, 0.5)) +
  labs(x = "Method", y = "Protected Area ID",
       color = "", 
       alpha = "")
```

4. Let's use zonal statistics to summarize human modification by access type!

```{r}

```
